\documentclass[9pt]{article}
\usepackage[left=1in, right=1in, bottom=1in, top=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{subfigure}

\title{PGS Line-Solver}

\begin{document}
\maketitle

Consider a stacked collection of rigid bodies in contact. In standard PGS, we loop through each contact constraint and update each constraint impulse sequentially, clamping each value to be positive on each iteration. 

An alternative approach would be to form ``lines'' of constraints, representing the most strongly coupled constraints, and apply a tridiagonal solve to this line of constraints at each PGS iteration. The impulses are clamped to be positive after each line solve.

To test this idea, consider a column of spheres, potentially with varying masses, dropped on the ground. Each sphere only has one degree of freedom, as we ignore lateral motion and rotation.

At each contact point between spheres $i$ and $i+1$, we have the following constraint between the sphere velocities:
\[
v_i - v_{i+1} = 0
\]
The Jacobian for each contact is therefore $J = [1\ \ {-1}]$.

The system to solve is given by 
\[
JM^{-1}J^T \lambda = \eta
\]
where 
\[
JM^{-1}J^T = 
\left[
\begin{array}{cccccc}
\frac{1}{m_g} + \frac{1}{m_1} & -\frac{1}{m_1} & 0 & \hdots & \hdots & 0 \\
-\frac{1}{m_1} & \frac{1}{m_1} + \frac{1}{m_2} & -\frac{1}{m_1} & 0 & \hdots & 0 \\
0 & -\frac{1}{m_2} & \frac{1}{m_2} + \frac{1}{m_3} & -\frac{1}{m_2} & 0 & 0 \\
\vdots & 0 & \ddots & \ddots & \ddots & 0 \\
0 & \hdots & 0 & -\frac{1}{m_{N-1}} & \frac{1}{m_{N-1}} + \frac{1}{m_N} & -\frac{1}{m_{N-1}} \\
\end{array} 
\right]
\]
and 
\[
\eta = \frac{1}{\Delta t}\zeta - J\left(\frac{1}{\Delta t}v^1 + M^{-1}f_{ext}\right)
\]

The third constraint is the contact constraint between the wheel and ground, and is only active if the wheel and ground are in contact:
\[
(\vec{u}_w + \vec{\omega}_w \times \vec{r}_c) \cdot \hat{n}_g = 0
\]
Since the wheel is a perfect circle, the contact vector $\vec{r}_c$ is always orthogonal to the ground normal, so $(\vec{\omega}_w \times \vec{r}_c) \cdot \hat{n}_g = 0$. Since the contact impulse can only push the wheel away, we make sure that the corresponding component of $\lambda$ is greater than zero.

The final constraint is the friction constraint: 
\[
(\vec{u}_w + \vec{\omega}_w \times \vec{r}_c) \cdot \hat{\tau}_g = 0
\]
The component of $\lambda$ corresponding to this constraint is bounded by the normal impulse multiplied by the coefficient of friction, but that is ignored for simplicity.

We first solve for the constraint impulses $\lambda$ by solving the system:
\[
J V_1 = J (V_0 + M^{-1} F_{ext} \Delta t) + J M^{-1} J^T \lambda
\]
and then update the velocities:
\[
V_1 = V_0 + M^{-1} F_{ext} \Delta t + M^{-1} J^T \lambda
\]
The inverse mass matrix is 
\[
M^{-1} = 
\left[
\begin{array}{cccccc}
\frac{1}{m_b} & 0 & 0 & 0 & 0 & 0 \\
0 & \frac{1}{m_b} & 0 & 0 & 0 & 0 \\
0 & 0 & \frac{1}{I_b} & 0 & 0 & 0 \\
0 & 0 & 0 & \frac{1}{m_w} & 0 & 0 \\
0 & 0 & 0 & 0 & \frac{1}{m_w} & 0 \\
0 & 0 & 0 & 0 & 0 & \frac{1}{I_w} \\
\end{array} 
\right]
\]
The forcing vector is 
\[
F_{ext} = 
\left[
\begin{array}{c}
f_{sp,x} \\
f_{sp,y} - m_b g \\
0 \\
0 \\
-m_w g \\
0
\end{array} 
\right]
\]
All components of $JV_1$ are zero except where we have elastic collisions or a specified motion, such as that of the bar/socket system. For this constraint, we have that 
\[
JV_1^{c0} = 
\left[
\begin{array}{c}
u_s(t + \Delta t)
\end{array} 
\right]
\]
We can add Baumgarte stabilization to the bar/wheel constraint by add a term that penalizes constraint violations:
\[
JV_1^{c1} = 
\left[
\begin{array}{c}
-\beta \left[x_w - \left(x_b + \frac{L}{2}\sin \theta_b\right)\right] \\
-\beta \left[y_w - \left(y_b - \frac{L}{2}\cos \theta_b\right)\right]
\end{array} 
\right]
\]
We can also add Baumgarte stabilization to the wheel contact constraint by adding a term proportional to the penetration depth of the wheel into the ground $d_{wg}$:
\[
JV_1^{c2} = 
\left[
\begin{array}{c}
-\beta d_{wg}
\end{array} 
\right]
\]

TODO elastic collision can be handled by not setting $JV_1$ to zero

\end{document}

